---
title: "Spearman's Correlation Analysis"
author: "Nandhini"
output: html_document
---
Showing the correlation analysis of early-stage patients, the same was used for Lung tissue and Late-stage patients
- Correlation between Immune infiltrates and Immune infiltrates in tumours and Lung tissue
- Correlation between Immune infiltrate and T-cell functionality 
- Correlation between Immune infiltrates and T-cell differentiation subsets 

```{r setup, include=FALSE}

# importing libraries #
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(readxl)
library(Hmisc)
library(tidyr)
library(RColorBrewer)

# Set WD #
knitr::opts_knit$set("~/xxx/xxxx/xxxx/xxxx")
```

1) Input the data table with percentages of immune infiltrates 
```{r Data Input}

#loading early-stage patients
early_data = read_excel("~/xxxx/xxxx/xxxx/NSCLC_Early_and_late_Immune_percentages.xlsx")
early_data = as.data.frame(early_data)

```

2) Correlation for Early-stage patients
```{r Correlation Analysis}

### Correlation heatmap ## 
## Spearman's Correlation 
rownames(early_data) = early_data$Patients
early_data = early_data[,-1]
early_data = as.matrix(early_data)
early_data = apply(early_data, 2, as.numeric) ## make sure they are numeric
early_corr = rcorr(early_data, type = "spearman")


# Function for correlation heatmap
# converting p,n, r values of correlation results into dataframes
# correlation coefficients
data.frame(early_corr$r) %>% head() %>% knitr::kable() ## use library(knitr for kable())
# p values; significanes
data.frame(early_corr$p) %>% head(n=5) %>% knitr::kable()
# n; number of patients
data.frame(early_corr$n) %>% head(n=3) %>% knitr::kable()

# function to convert r,p and n  values as list to dataframes
cors = function(df) { 
   # turn all three matrices (r, n, and P into a dataframe)
   M = Hmisc::rcorr(as.matrix(df), type = "spearman")
   # return the three data frames in a list return(Mdf)
   Mdf = map(M, ~data.frame(.x))
}

# Implementing the cors function
cors(early_data) %>%
 map(~rownames_to_column(.x, var="measure1")) %>%
 # format each data set (r,P,n) long 
 map(~pivot_longer(.x, -measure1, "measure2")) %>%
 # merge our three list elements by binding the rows
 bind_rows(.id = "id") %>%
 pivot_wider(names_from = id, values_from = value) %>%
 mutate(sig_p = ifelse(P < .05, T, F), p_if_sig = ifelse(P <.05, P, NA), r_if_sig = ifelse(r <.05, r, NA)) %>% 
 head() %>%
 knitr::kable()

formatted_cors <- function(df){
 cors(df) %>%
 map(~rownames_to_column(.x, var="measure1")) %>%
 map(~pivot_longer(.x, -measure1, "measure2")) %>% 
 bind_rows(.id = "id") %>%
 pivot_wider(names_from = id, values_from = value) %>%
 mutate(sig_p = ifelse(P < .05, T, F), p_if_sig = ifelse(P <.05, P, NA), r_if_sig = ifelse(P <.05, r, NA)) 
}

formatted_cors(early_data) %>% head() %>% knitr::kable()

## Applying the cor function to early-stage patients 
early_cor_df = formatted_cors(early_data)

##saving the correlation results
write.csv(early_cor_df, "~/xxxx/xxxx/NSCLC_early_correlation_plots/NSCLC_correlation_early.csv")


```

## 3) Visualization
```{r}

# making pretty heatmaps
# 3.1) Correlation between Immune cells vs Immune cells

# filtered heatmaps
early_cor_im_df = 
   early_cor_df %>% 
  filter(
    measure1 %in% c("CD8 T","Tconv","Tregs","B cells","Neutrophils","Monocytes","DCs","NKT","NK"),
    measure2 %in% c("CD8 T","Tconv","Tregs","B cells","Neutrophils","Monocytes","DCs","NKT","NK"))

# adding significance values 
early_cor_im_df$sig_symbol = ifelse(early_cor_im_df$p_if_sig <= 0.0005, "***",
                            ifelse(early_cor_im_df$p_if_sig <= 0.005, "**",
                            ifelse(early_cor_im_df$p_if_sig <= 0.05, "*", NA)))

# Reorder the levels
early_cor_im_df$measure1 = as.factor(early_cor_im_df$measure1)
early_cor_im_df$measure1 = factor(early_cor_im_df$measure1, levels = c("CD8 T","Tconv","Tregs","B cells","NK","NKT","Neutrophils","Monocytes","DCs"))

early_cor_im_df$measure2 = as.factor(early_cor_im_df$measure2)
early_cor_im_df$measure2 = factor(early_cor_im_df$measure2, levels = c("DCs","Monocytes","Neutrophils","NKT","NK","B cells","Tregs","Tconv","CD8 T"))

# Replacing 1s in the diagonal by NA 
early_cor_im_df[early_cor_im_df == 1] = NA

# Correlation heatmap: Immune cells vs Immune cells
ggplot(data = early_cor_im_df, aes(x = measure1, y = measure2, fill = r)) +
  geom_raster(data = early_cor_im_df, aes(x = measure1, y = measure2, fill = r)) +
  geom_tile(color = "white")+
  geom_text(data = early_cor_im_df, aes(x = measure1, y = measure2, label = round(r, digits=2))) +
  geom_text(data = early_cor_im_df, aes(x = measure1, y = measure2, label = sig_symbol),
            nudge_y = 0.25, nudge_x = 0.25, size = 5) +
  scale_fill_gradient2(mid="#ffffff",low="#053061",high="#67001f", limits=c(-0.6,0.6)) +
    theme(panel.background = element_blank(),axis.text.x = element_text(angle = 60, hjust=1))+
  labs(x = NULL, y = NULL, fill = "Spearman's\nCorrelation", title="Immune cells vs Immune cells")


## 3.2) Correlation between Immune cells vs T cell functionality 
# Filter immune cells and T-cell functionality
early_cor_fun_df =
  early_cor_df %>%
  filter(
    measure2 %in% c ("%CD137_CD4", "%CD137_CD8","%Cytokines_CD4","%Cytokines_CD8"),
    measure1 %in% c("CD8 T","Tconv","Tregs","B cells","Neutrophils","Monocytes","DCs","NKT","NK"))

# Adding significance values
early_cor_fun_df$sig_symbol = ifelse(early_cor_fun_df$p_if_sig <= 0.0005, "***",
                            ifelse(early_cor_fun_df$p_if_sig <= 0.005, "**",
                            ifelse(early_cor_fun_df$p_if_sig <= 0.05, "*", NA)))

# Reorder the levels
early_cor_fun_df$measure1 = as.factor(early_cor_fun_df$measure1)
early_cor_fun_df$measure1 = factor(early_cor_fun_df$measure1, levels = c("DCs","Monocytes","Neutrophils","NKT","NK","B cells","Tregs","Tconv","CD8 T"))

early_cor_fun_df$measure2 = as.factor(early_cor_fun_df$measure2)
early_cor_fun_df$measure2 <- factor(early_cor_fun_df$measure2, levels = c("%CD137_CD4","%CD137_CD8","%Cytokines_CD4","%Cytokines_CD8"))

# Correlation Heatmap: Immune cells vs T cell functionality
ggplot(data = early_cor_fun_df, aes(x = measure2, y = measure1, fill = r)) +
  geom_raster(data = early_cor_fun_df, aes(x = measure2, y = measure1, fill = r)) +
  geom_tile(color='white') +
  geom_text(data = early_cor_fun_df, aes(x = measure2, y = measure1, label = round(r, digits=2),size = 2)) +
  geom_text(data = early_cor_fun_df, aes(x = measure2, y = measure1, label = sig_symbol),
            nudge_y = 0.25, nudge_x = 0.25, size = 4.5) +
  scale_fill_gradient2(mid="#ffffff",low="#053061",high="#67001f", limits=c(-0.6,0.6)) +
    theme(panel.background = element_blank(),axis.text.x = element_text(angle = 60, hjust=1))+
  labs(x = NULL, y = NULL, fill = "Spearman's\nCorrelation", title="Immune cells vs T cell functionality")


## 3.3) Correlation between Immune cells vs T cell differentiation subsets
# Filter immune cells and T cell differentiation subsets
early_cor_dif_df = 
   early_cor %>% 
  filter(
    measure2 %in% c("CD8 T","Tconv","Tregs","B cells","Neutrophils","Monocytes","DCs","NKT","NK"),
    measure1 %in% c("CD8_TCM","CD8_TEM","CD8_TEMRA","CD8_TN",
                    "CD4_TCM","CD4_TEM","CD4_TEMRA","CD4_TN",
                    "Tregs_TCM","Tregs_TEM","Tregs_TEMRA","Tregs_TN"))

# Adding significance values 
early_cor_dif_df$sig_symbol = ifelse(early_cor_dif_df$p_if_sig <= 0.0005, "***",
                            ifelse(early_cor_dif_df$p_if_sig <= 0.005, "**",
                            ifelse(early_cor_dif_df$p_if_sig <= 0.05, "*", NA)))

# Reorder the levels
early_cor_dif_df$measure2 = as.factor(early_cor_dif_df$measure2)
early_cor_dif_df$measure2 <- factor(early_cor_dif_df$measure2, levels = c("DCs","Monocytes","Neutrophils","NKT","NK","B cells","Tregs","Tconv","CD8 T"))

# Correlation Heatmap; Immune cells vs T cell differentiation subsets
ggplot(data = early_cor_dif_df, aes(x = measure1, y = measure2, fill = r)) +
  geom_raster(data = early_cor_dif_df, aes(x = measure1, y = measure2, fill = r)) +
  geom_tile(color='white') +
  geom_text(data = early_cor_dif_df, aes(x = measure1, y = measure2, label = round(r, digits=2))) +
  geom_text(data = early_cor_dif_df, aes(x = measure1, y = measure2, label = sig_symbol),
            nudge_y = 0.25, nudge_x = 0.25, size = 4.5) +
  scale_fill_gradient2(mid="#ffffff",low="#053061",high="#67001f", limits=c(-0.65,0.65)) +
    theme(panel.background = element_blank(),axis.text.x = element_text(angle = 60, hjust=1))+
  labs(x = NULL, y = NULL, fill = "Spearman's\nCorrelation", title="Immune cells vs T cell differentiation subsets")


```

